---
title: "Exercise 4_1: OSM Data"
subtitle: 'Introduction to Geospatial Techniques for Social Scientists in R'
author: Stefan Jünger & Dennis Abel
format: 
  html:
    embed-resources: true
    code-copy: true
execute:
  echo: true
editor: visual
callout-icon: false
callout-appearance: minimal
---

### Exercises

::: callout-note
## `r emo::ji("weight_lifting_man")` Exercise 1

Pick a city of the world to your liking and retrieve its bounding box. Then plot it: What do you see?
:::

::: {.callout-caution collapse="true"}
## `r emo::ji("bulb")` Tip

The function you may search for is part of the `osmdata` package. Have a look at the slides.
:::

::: callout-note
## `r emo::ji("weight_lifting_man")` Exercise 2

Please choose a couple of building types you are interested in and set them as key and value pairs. You can find a list of building types in the [Overpass API documentation](https://wiki.openstreetmap.org/wiki/Key:building). But don't forget to set the timeout query parameters using the `osmdata::opq()` function first.
:::

::: {.callout-caution collapse="true"}
## `r emo::ji("bulb")` Tip

First, specify the bounding box like before; second, the query parameters; and third, the key and value pairs. Try using a pipe workflow, as it makes it easier.
:::

::: callout-note
## `r emo::ji("weight_lifting_man")` Exercise 3

Download the data using the `osmdata::osmdata_sf()` function and extract only the polygons.
:::

::: {.callout-caution collapse="true"}
## `r emo::ji("bulb")` Tip

The downloaded data is a list. The polygons are a named list element that you can extract with its name `osm_polygons`, just like a variable in a data table.
:::

::: callout-note
## `r emo::ji("weight_lifting_man")` Exercise 4

Take some time to browse through the data. Depending on your chosen building type, you may find some interesting information. You can also plot the data you have just downloaded.
:::

::: {.callout-caution collapse="true"}
## `r emo::ji("bulb")` Tip

You may consider converting the data into a tibble using `tibble::as_tibble()` and maybe a `sf::st_as_sf()` afterward for a nicer browsing experience.
:::

### Solutions

::: {.callout-tip collapse="true"}
## `r emo::ji("white_check_mark")` Solution 1

```{r}
boothbay <- osmdata::getbb("Boothbay")

plot(boothbay)

# We see two points, which build the extremes of the bounding box.
```
:::

::: {.callout-tip collapse="true"}
## `r emo::ji("white_check_mark")` Solution 2

```{r}
boothbay <-
  osmdata::getbb("Boothbay") |>  
  osmdata::opq(timeout = 25*100) |> 
  osmdata::add_osm_feature(
    key = "building", 
    value = c("house", "residential", "apartments", "detached", "bungalow")
  )
```
:::

::: {.callout-tip collapse="true"}
## `r emo::ji("white_check_mark")` Solution 3

```{r}
boothbay <-
  osmdata::getbb("Boothbay") |>  
  osmdata::opq(timeout = 25*100) |> 
  osmdata::add_osm_feature(
    key = "building", 
    value = c("house", "residential", "apartments", "detached", "bungalow")
  ) |> 
  osmdata::osmdata_sf()

boothbay <- boothbay$osm_polygons
```
:::

::: {.callout-tip collapse="true"}
## `r emo::ji("white_check_mark")` Solution 4

```{r}
library(tmap)

boothbay <-
  boothbay |> 
  tibble::as_tibble() |> 
  sf::st_as_sf()

boothbay

# tmap_mode(("view"))

tm_shape(boothbay) +
  tm_polygons()
```
:::
